---
title: "MSDS 6370 Sampling Design"
output: html_notebook
---

### Load the data

Load necessary libraries

```{r}
require(survey)
require(dplyr)
```

Pulling in data only if RDS file has not already been created.

```{r}
setwd("../Analysis/Data")
if(!file.exists("popData.rds")){
    pop.data <- NULL
    for(i in 1:length(list.files())){
        data.frag <- read.csv(paste("dataset",i,".csv",sep = ""), header = FALSE)
        pop.data <- rbind(pop.data, data.frag)
    }
    colnames(pop.data) <- c("tripduration",
                            "starttime",
                            "stoptime",
                            "start_station_id",
                            "start_station_name",
                            "start_station_latitude",
                            "start_station_longitude",
                            "end_station_id",
                            "end_station_name",
                            "end_station_latitude",
                            "end_station_longitude",
                            "bikeid",
                            "usertype",
                            "birth year",
                            "gender",
                            "LinearDistance",
                            "DayOfWeek",
                            "TimeOfDay",
                            "HolidayFlag",
                            "PRCP",
                            "SNOW",
                            "TAVE",
                            "TMAX",
                            "TMIN")
    saveRDS(pop.data, "popData.rds")
    rm(data.frag)
} else pop.data <- readRDS("popData.rds")
```

### Task 1
#### Stratified Sampling (Proportional Allocation)


Compute the sample size for SRS with the 2 step calculation from lab 5 - Q1 - check for FPC adjustment (sample size > 10% population)
```{r}
####Compute the sample size for a SRS
##step 1


```

Compute Estimate using survey library with SRS and survey design of SRS 
```{r}


```

Compute Estimate using survey library with SRS, but use surveyDesign of Stratified Sample
```{r}


```

Compute the Design Effect (SE stratified design / SE SRS Design)
```{r}


```

Compute Sample Size for Stratified Design Effect Modification (see 5.6 slides)
```{r}


```


Compute Proportional Allocation sample sizes amongst stratum
```{r}

sampleSize = 250000

pop.data.TOD <- pop.data[c("tripduration", "DayOfWeek")]
head(pop.data.TOD)

  # Identify Frequency of DayOfWeek Stratum
ToDFreq <- as.data.frame(table(pop.data.TOD$DayOfWeek))
names(ToDFreq)[1] = 'DayOfWeek'
ToDFreq


ToDFreq$N = nrow(pop.data.TOD)
ToDFreq$p = ToDFreq$Freq/ToDFreq$N
ToDFreq$SampSizeh = ToDFreq$p * sampleSize
ToDFreq$SampSizehRounded = round(ToDFreq$SampSizeh)

print(sum(ToDFreq$SampSizeh))

head(ToDFreq)
tail(ToDFreq)


```

Utilize Sample Sizes to compute an Estimate of the mean, using Stratified Design with appropriate sample size
```{r}
  
pop.data.TOD <- pop.data[c("tripduration", "DayOfWeek")]
pop.data.TOD

pop.data.TOD$N <- NA

  # Identify Frequency of DayOfWeek Stratum
ToDFreq <- as.data.frame(table(pop.data.TOD$DayOfWeek))
names(ToDFreq)[1] = 'DayOfWeek'
ToDFreq

#pop.data.TOD$Nh[pop.data.TOD$DayOfWeek == "Friday"] <- 829007
#pop.data.TOD$Nh[pop.data.TOD$DayOfWeek == "Monday"] <- 806725
#pop.data.TOD$Nh[pop.data.TOD$DayOfWeek == "Saturday"] <- 734053
#pop.data.TOD$Nh[pop.data.TOD$DayOfWeek == "Sunday"] <- 674524
#pop.data.TOD$Nh[pop.data.TOD$DayOfWeek == "Thursday"] <- 820587
#pop.data.TOD$Nh[pop.data.TOD$DayOfWeek == "Tuesday"] <- 839970
#pop.data.TOD$Nh[pop.data.TOD$DayOfWeek == "Wednesday"] <- 857450

print(nrow(pop.data.TOD))

pop.data.TOD$N[pop.data.TOD$DayOfWeek == "Friday"] <- 829007
pop.data.TOD$N[pop.data.TOD$DayOfWeek == "Monday"] <- 806725
pop.data.TOD$N[pop.data.TOD$DayOfWeek == "Saturday"] <- 734053
pop.data.TOD$N[pop.data.TOD$DayOfWeek == "Sunday"] <- 674524
pop.data.TOD$N[pop.data.TOD$DayOfWeek == "Thursday"] <- 820587
pop.data.TOD$N[pop.data.TOD$DayOfWeek == "Tuesday"] <- 839970
pop.data.TOD$N[pop.data.TOD$DayOfWeek == "Wednesday"] <- 857450

tail(pop.data.TOD)

  # Create SurveyDesign
mydesign <- svydesign(id = ~1, strata = ~DayOfWeek, data = pop.data.TOD, fpc = ~N)

#mydesign <- postStratify(design = mydesign, strata = ~DayOfWeek, population = ToDFreq)

svymean(~tripduration, design = mydesign)
```


#### Neyman Sampling Design

```{r}
  # Identify Frequency of DayOfWeek Stratum
ToDFreq <- as.data.frame(table(pop.data.TOD$DayOfWeek))
names(ToDFreq)[1] = 'DayOfWeek'

stdDevs <- tapply(pop.data.TOD$tripduration, pop.data.TOD$DayOfWeek, sd)
NhSh <- ToDFreq$Freq * stdDevs
NhSh.ratio <- NhSh/sum(NhSh)
sampsRaw <- round(n0srs)*NhSh.ratio
Neyman.samples <- round(sampsRaw)

cbind(ToDFreq, stdDevs, NhSh, NhSh.ratio, sampsRaw, Neyman.samples)
```
<br>
By rounding the calculated strata sample counts in our table output above, we now have definitive Neyman sample sizes for each stratum. To help validate these numbers' derivation, we choose to calculate their sum - their sum should calculate to our n<sub>0,srs</sub> value.
```{r}
sum(Neyman.samples)
```
<br>
As observed above, our sum is just one sample short of our n<sub>0,srs</sub> value. This is due to rounding error when deriving our whole number sample sizes from our raw, calculated sample sizes. For this reason, we choose to select the largest raw sample size which contains a decimal value less than 0.5 and force it to round up. In this case, Saturday's value of 29459.48 fits this criterion. Therefore, we manually assign a value of 29460 to replace its rounded value of 29460.
```{r}
Neyman.samples[Neyman.samples %in% 29459] <- 29460
sum(Neyman.samples)
```

```{r}
pop.data.Neyman <- data.frame(DayOfWeek = pop.data.TOD$DayOfWeek, tripduration = pop.data.TOD$tripduration)
pop.data.Neyman$N <- NA

for(i in seq(1,7)) pop.data.Neyman$N[pop.data.Neyman$DayOfWeek == attributes(Neyman.samples[i])] <- Neyman.samples[i]

head(pop.data.Neyman)
tail(pop.data.Neyman)
```
```{r}
prop.samp <- function(i) {
    sample_n(pop.data.Neyman[pop.data.Neyman$DayOfWeek == attributes(Neyman.samples[i]),], Neyman.samples[i])
}

Friday.samples <- prop.samp(1)
Monday.samples <- prop.samp(2)
Saturday.samples <- prop.samp(3)
Sunday.samples <- prop.samp(4)
Thursday.samples <- prop.samp(5)
Tuesday.samples <- prop.samp(6)
Wednesday.samples <- prop.samp(7)

NeySamp <- rbind(Sunday.samples,
                 Monday.samples,
                 Tuesday.samples,
                 Wednesday.samples,
                 Thursday.samples,
                 Friday.samples,
                 Saturday.samples)
```

```{r}
NeySamp <- merge(NeySamp, ToDFreq, by = "DayOfWeek")
```


```{r}
  # Create SurveyDesign
mydesign <- svydesign(id = ~1, strata = ~DayOfWeek, data = NeySamp) # Dropped since no correction needed: , fpc = ~Freq)

#mydesign <- postStratify(design = mydesign, strata = ~DayOfWeek, population = ToDFreq)

svymean(~tripduration, design = mydesign)
```

```{r}
confint(svymean(~tripduration, design = mydesign))
```


